---
title: "Covid2"
author: "Briody"
date: "3/25/2020"
output: html_document
---


```{r setup, include=FALSE}
#install.packages("showtext")
#install.packages("plotly")
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(fig.width=8, fig.height=5) 
library(tidyverse)
library(plotly)
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#0072B2", "#D55E00", "#CC79A7", "#999999")
#"#F0E442" replaced yellow last with grey "#999999"
theme_set(theme_minimal())
#covid <- readr::read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv")
#covid[covid$`3/18/20` == 162, '3/18/20'] <- 288 #replace wrong number
```

# Import choices
## Johns Hopkins
On March 23, Johns Hopkins reorganized their data files halting the usefulness of the original Covid file. Two new sources were promised. As of 3/24/20 only one time series has been delivered (confirmed and deaths by country), 

time_series_covid19_confirmed_global.csv
time_series_covid19_deaths_global.csv

The county (and by summing, state) numbers starting 3/22/20 are available but in a daily update and not a single file.

03-22-2020.csv	inside csse_covid_19_daily_reports

```{r}
covid2jh <- readr::read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")

covid2jh_long <- gather(covid2jh, date, confirmed, contains("/20"))
covid2jh_long$date <- as.Date(covid2jh_long$date, "%m/%d/%y")
```

### Europe Test
```{r}
europe <- c("France", "Germany", "Italy",  "Poland", "Spain",  "Switzerland", "United Kingdom")
#covid2jh_euro <- filter(covid2jh_long, `Country/Region` %in% europe & `Province/State` %in% c(europe,NA))
covid2jh_euro <- filter(covid2jh_long, `Country/Region` %in% europe & `Province/State` %in% c(europe,NA), date > as.Date('2020-01-25'))
covid2jh_euro
```

Plot (Regular Scale)

```{r}
p <- ggplot(covid2jh_euro, aes(x = date, y = confirmed)) + 
  geom_point(aes(shape = covid2jh_euro$`Country/Region`, color = covid2jh_euro$`Country/Region`)) + 
  labs(x="", y = "Confirmed", shape = "Country", 
       title = "Confirmed COVID-19 Cases", 
       subtitle = "Selected European Countries", 
       caption = "Source: CDC and Johns Hopkins via GitHub") +
  scale_color_manual(name = " ",
                     labels = europe, 
                     values = cbPalette) +
  scale_shape_manual(name = " ",
                     labels = europe, 
                     values = c(15, 16, 17, 3, 7, 8, 11))
p + scale_x_date(date_labels = "%b%d")
```

Plot (Log Scale)
```{r}
p <- ggplot(covid2jh_euro, aes(x = date, y = confirmed)) + 
  geom_point(aes(shape = covid2jh_euro$`Country/Region`, color = covid2jh_euro$`Country/Region`)) + 
  labs(x="", y = "Confirmed", shape = "Country", 
       title = "Confirmed COVID-19 Cases on a Log Scale", 
       subtitle = "Selected European Countries", 
       caption = "Source: CDC and Johns Hopkins via GitHub") + 
  scale_color_manual(name = " ",
                     labels = europe, 
                     values = cbPalette) +
  scale_shape_manual(name = " ",
                     labels = europe, 
                     values = c(15, 16, 17, 3, 7, 8, 11))
p + scale_x_date(date_labels = "%b%d") + scale_y_continuous(trans = 'log10', labels = scales::comma)
```


## Corona Data Scraper SoSo - missing some days USA Data seems better for state and county totals
Others report using Corona Data Scraper (https://coronadatascraper.com/#home) 

csv (tidy format)
https://coronadatascraper.com/timeseries-tidy.csv

```{r}
covid2cds <- readr::read_csv("https://coronadatascraper.com/timeseries-jhu.csv")
problems(covid2cds)
#view(covid2cds)
```

```{r}
covid2cds_long <- gather(covid2cds, date, confirmed, contains("2020-"))
covid2cds_long$date <- as.Date(covid2cds_long$date, "%Y-%m-%d")
```

### Cook County Plot
```{r}
(covid2cds_cook <- filter(covid2cds_long, county == "Cook County" & state == "IL"))
p <- ggplot(covid2cds_cook, aes(x = date, y = confirmed)) + 
  geom_point() + 
  xlab("") + 
  scale_x_date(date_labels = "%b%d") 
p 
```

### State Totals
```{r}
# Warning: NT has city totals as well.
# Must filter out state = NY and population > 10M
us_only <- filter(covid2cds_long, country == "USA")
(by_state <- group_by(us_only, state, date))
(per_day <- summarise(by_state, state_total = sum(confirmed, na.rm = TRUE)))
```

#### Get Illinois
```{r}
(illinois <- filter(per_day, state == "IL"))
p <- ggplot(illinois, aes(x = date, y = state_total)) + 
  geom_point() + 
  xlab("") + 
  scale_x_date(date_labels = "%b%d") 
p 
```


## USA Facts _ use this instead of coronadatascraper
or usafacts (https://usafacts.org/visualizations/coronavirus-covid-19-spread-map/)

from webpage
https://static.usafacts.org/public/data/covid-19/covid_confirmed_usafacts.csv?_ga=2.195427820.1316833666.1585136621-915307346.1585136621

https://static.usafacts.org/public/data/covid-19/covid_deaths_usafacts.csv?_ga=2.195427820.1316833666.1585136621-915307346.1585136621

links in GitHub comment
 use first below
https://static.usafacts.org/public/data/covid-19/covid_confirmed_usafacts.csv
https://static.usafacts.org/public/data/covid-19/covid_deaths_usafacts.csv?_ga=2.210670165.1998648927.1584971390-1836018421.1584457034

```{r}
#recreate illinois plot from above the add other states
covid2usafacts <- readr::read_csv("https://static.usafacts.org/public/data/covid-19/covid_confirmed_usafacts.csv")
problems(covid2usafacts)
covid2usafacts_long <- gather(covid2usafacts, date, confirmed, contains("/20"))
covid2usafacts_long$date <- as.Date(covid2usafacts_long$date, "%m/%d/%y")
```

### Cook County
```{r}
(cook <- filter(covid2usafacts_long, State == "IL" & `County Name` == "Cook County"))
p <-  ggplot(cook, aes(x = date, y = confirmed)) + 
  geom_point(color = "Blue") + 
  xlab("") + 
  scale_x_date(date_labels = "%b%d") + 
  labs(x="", y = "Confirmed", 
       title = "Confirmed COVID-19 Cases in Cook County", 
       caption = "Source: USA Facts (usafacts.org)")
p
ggplotly(p)
```
 ### Cook County on Log Scale
```{r}
q <- p  + 
  scale_y_continuous(trans = 'log10', labels = scales::comma) + 
  labs(title = "Confirmed COVID-19 Cases in Cook County", 
       subtitle = "Plotted on a Log Scale", 
       caption = "Source: USA Facts (usafacts.org)")
q
ggplotly(q)
```
 
 
 ### Cook County Percent Increase
```{r}
cook2 <- cook %>% 
  mutate(pct_change = (confirmed/lag(confirmed)-1) *100) %>% 
  filter(date > "2020-03-01")
cook2
ggplot(cook2, aes(x = date, y = pct_change)) + 
  geom_point( color = "Blue") + 
  xlab("") + 
  scale_x_date(date_labels = "%b%d") + 
  labs(x="", y = "% Increase in Confirmed", 
       title = "Confirmed COVID-19 Cases in Cook County",
       subtitle = "Daily Percent Increase",
       caption = "Source: USA Facts (usafacts.org)")
```

### Illinois
```{r}
illinois <- filter(covid2usafacts_long, State == "IL") %>% 
  group_by(date) %>% 
  summarise(total = sum(confirmed))
illinois
```

create plot
```{r}
p <-  ggplot(illinois, aes(x = date, y = total)) + 
  geom_point(color = "Blue") + 
  xlab("") + 
  scale_x_date(date_labels = "%b%d") + 
  labs(x="", y = "Confirmed", 
       title = "Confirmed COVID-19 Cases in Illinois County", 
       caption = "Source: USA Facts (usafacts.org)")
p
ggplotly(p)
```

Illinois on log scale:
```{r}
```{r}
q <- p  + 
  scale_y_continuous(trans = 'log10', labels = scales::comma) + 
  labs(title = "Confirmed COVID-19 Cases in Illinois", 
       subtitle = "Plotted on a Log Scale", 
       caption = "Source: USA Facts (usafacts.org)")
q
ggplotly(q)
```

Illinois Percent Increase
```{r}
illinois2 <- illinois %>% 
  mutate(pct_change = (total/lag(total)-1) *100) %>% 
  filter(date > "2020-03-01")
illinois2
ggplot(illinois2, aes(x = date, y = pct_change)) + 
  geom_point( color = "Blue") + 
  xlab("") + 
  scale_x_date(date_labels = "%b%d") + 
  labs(x="", y = "% Increase in Confirmed", 
       title = "Confirmed COVID-19 Cases in Illinois",
       subtitle = "Daily Percent Increase",
       caption = "Source: USA Facts (usafacts.org)")
```

### Selected States
Filtering states
```{r}
state_list <- c("IL", "WA", "NJ", "LA", "NY")
states <- filter(covid2usafacts_long, State %in% state_list & date >= "2020-03-01") %>% 
  group_by(State,date) %>% 
  summarise(total = sum(confirmed))  %>% 
  mutate(pct_change = (total/lag(total)-1) *100)
states
```

Making the plot
```{r}
states_plot <- ggplot(states, aes(x= date, y = total, color = State)) +
  geom_point() +
  geom_path() +
  scale_y_continuous(trans = "log10")

states_plot
```

Find percent increase
```{r}
states_percent_data <- filter(states, date > "2020-03-10")

states_percent_plot <- ggplot(states_percent_data, aes(x = date, y = pct_change, color = State)) +
  geom_point() + 
  geom_path()

states_percent_plot
ggplotly(states_percent_plot)
```



## JHU Manual download
```{r}
covid2_jh_manal <- readr::read_csv("covid.csv")
problems(covid2_jh_manal)
```
### Illinois Starts March 22
```{r}
(illinois_manual <- filter(covid2_jh_manal, Province_State == "Illinois"))

```


### Cook County Starts March 22
```{r}
(cook_manual <- filter(covid2_jh_manal, Combined_Key == "Cook, Illinois, US"))
```

