---
title: "Covid"
author: "Briody"
date: "3/14/2020"
output: html_document
---




```{r setup, include=FALSE}
#install.packages("showtext")
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(fig.width=8, fig.height=5) 
library(tidyverse)
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#0072B2", "#D55E00", "#CC79A7", "#999999")
#"#F0E442" replaced yellow last with grey "#999999"
theme_set(theme_minimal())
covid <- readr::read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv")
covid[covid$`3/18/20` == 162, '3/18/20'] <- 288 #replace wrong number
```

## Change format from data table to each line is a case.
```{r}
covid_long <- gather(covid, date, confirmed, contains("/20"))
#covid_long <- gather(covid, date, confirmed, `1/22/20`:`3/15/20`)
covid_long$date <- as.Date(covid_long$date, "%m/%d/%y")
#Below is after r is updated?
#covid %>% 
#  pivot_longer(c(`1/22/20`, `1/22/20`), names_to = "date", values_to = "diagnosed")
```

## Cook County 
```{r}
covid_cook <- filter(covid_long, `Province/State` == "Cook County, IL")
p <- ggplot(covid_cook, aes(x = date, y = confirmed)) + 
  geom_point() + 
  xlab("")
p + scale_x_date(date_labels = "%d") 
# csv file stopped updating Cook county numbers.
```


## Illinois
```{r}
covid_IL <- filter(covid_long, `Province/State` == "Illinois")
covid_IL
```

```{r}
p <- ggplot(covid_IL, aes(x = date, y = confirmed)) + 
  geom_point() + 
  xlab("") 
p + scale_x_date(date_labels = "%d") 
```

## California
```{r}
covid_CAL <- filter(covid_long, `Province/State` == "California")
covid_CAL
```

```{r}
p <- ggplot(covid_CAL, aes(x = date, y = confirmed)) + 
  geom_point() + 
  xlab("") 
p + scale_x_date(date_labels = "%d") 
```




## Why we need logs
First subset selected European countries.
```{r}
europe <- c("France", "Germany", "Italy",  "Poland", "Spain",  "Switzerland", "United Kingdom")
covid_euro <- filter(covid_long, `Country/Region` %in% europe & `Province/State` %in% c(europe,NA), date > as.Date('2020-01-25') & date < as.Date('2020-03-10'))
covid_euro
```

Next, plot:
```{r}
p <- ggplot(covid_euro, aes(x = date, y = confirmed)) + 
  geom_point(aes(shape = covid_euro$`Country/Region`, color = covid_euro$`Country/Region`)) + 
  labs(x="", y = "Confirmed", shape = "Country", 
       title = "Confirmed COVID-19 Cases", 
       subtitle = "Selected European Countries", 
       caption = "Source: CDC and Johns Hopkins via GitHub") +
  scale_color_manual(name = " ",
                     labels = europe, 
                     values = cbPalette) +
  scale_shape_manual(name = " ",
                     labels = europe, 
                     values = c(15, 16, 17, 3, 7, 8, 11))
p + scale_x_date(date_labels = "%b%d")
#ggsave("log1.png")
```

Now, log the values.
```{r}
covid_euro_log <- covid_euro %>% 
  mutate(log_confirmed = log(confirmed+1,10)) #the +1 avoids log(0) which doesn't exist
```
Plot log(confirmed) for each day.
```{r}
p <- ggplot(covid_euro_log, aes(x = date, y = log_confirmed)) + 
  geom_point(aes(shape = covid_euro$`Country/Region`, color = covid_euro$`Country/Region`)) + 
  labs(x="", y = "Log(Confirmed)", shape = "Country", 
       title = "Confirmed COVID-19 Cases", 
       subtitle = "Selected European Countries", 
       caption = "Source: CDC and Johns Hopkins via GitHub") +
  scale_color_manual(name = " ",
                     labels = europe, 
                     values = cbPalette) +
  scale_shape_manual(name = " ",
                     labels = europe, 
                     values = c(15, 16, 17, 3, 7, 8, 11))
p + scale_x_date(date_labels = "%b%d")
#ggsave("log2.png")
```


Change y-axis scale.
```{r warning=FALSE}
p <- ggplot(covid_euro, aes(x = date, y = confirmed)) + 
  geom_point(aes(shape = covid_euro$`Country/Region`, color = covid_euro$`Country/Region`)) + 
  labs(x="", y = "Confirmed", shape = "Country", 
       title = "Confirmed COVID-19 Cases", 
       subtitle = "Selected European Countries", 
       caption = "Source: CDC and Johns Hopkins via GitHub") + 
  scale_color_manual(name = " ",
                     labels = europe, 
                     values = cbPalette) +
  scale_shape_manual(name = " ",
                     labels = europe, 
                     values = c(15, 16, 17, 3, 7, 8, 11))
p + scale_x_date(date_labels = "%b%d") + scale_y_continuous(trans = 'log10')
#ggsave("log3.png")
```

## Europe to date
```{r}
europe <- c("France", "Germany", "Italy",  "Poland", "Spain",  "Switzerland", "United Kingdom")
covid_euro_todate <- filter(covid_long, `Country/Region` %in% europe & `Province/State` %in% c(europe,NA), date > as.Date('2020-01-25'))
covid_euro_todate
```
 Europe to date plot
```{r}
p <- ggplot(covid_euro_todate, aes(x = date, y = confirmed)) + 
  geom_point(aes(shape = covid_euro_todate$`Country/Region`, color = covid_euro_todate$`Country/Region`)) + 
  labs(x="", y = "Confirmed", shape = "Country", 
       title = "Confirmed COVID-19 Cases", 
       subtitle = "Selected European Countries", 
       caption = "Source: CDC and Johns Hopkins via GitHub") +
  scale_color_manual(name = " ",
                     labels = europe, 
                     values = cbPalette) +
  scale_shape_manual(name = " ",
                     labels = europe, 
                     values = c(15, 16, 17, 3, 7, 8, 11))
p + scale_x_date(date_labels = "%b%d")
```

Change y-axis scale for to-date.
```{r warning=FALSE}
p <- ggplot(covid_euro_todate, aes(x = date, y = confirmed)) + 
  geom_point(aes(shape = covid_euro_todate$`Country/Region`, color = covid_euro_todate$`Country/Region`)) + 
  labs(x="", y = "Confirmed", shape = "Country", 
       title = "Confirmed COVID-19 Cases on a Log Scale", 
       subtitle = "Selected European Countries", 
       caption = "Source: CDC and Johns Hopkins via GitHub") + 
  scale_color_manual(name = " ",
                     labels = europe, 
                     values = cbPalette) +
  scale_shape_manual(name = " ",
                     labels = europe, 
                     values = c(15, 16, 17, 3, 7, 8, 11))
p + scale_x_date(date_labels = "%b%d") + scale_y_continuous(trans = 'log10')
```

## Cases by state
```{r}
states <- c("Alabama","Alaska","Arizona","Arkansas","California","Colorado","Connecticut","Delaware","District of Columbia","Florida","Georgia","Hawaii","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire","New Jersey","New Mexico","New York","North Carolina","North Dakota","Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island","South Carolina","South Dakota","Tennessee","Texas","Utah","Vermont","Virginia","Washington","West Virginia","Wisconsin","Wyoming")
covid_states <- filter(covid_long, `Country/Region` == 'US', `Province/State` %in% states, date == max(date)) %>% 
  select(`Province/State`,confirmed) %>% 
  arrange(desc(confirmed))
covid_states
```

## Time plot for top eight states
Get list:
```{r}
top_states_list <- covid_states %>% 
  select(`Province/State`) %>% 
  head(8)
top_states_list
```

Use only data from top states via join
```{r}
top_states <- covid_long %>% 
  inner_join(top_states_list, by = "Province/State")
top_states
```

```{r}
#BAD _ DON"T USE
top_states <- inner_join(covid_long, top_states_list)
top_states
```


First confirmed for top states was March 10
```{r}
top_states_recent <-  top_states %>% 
  filter(date > as.Date('2020-03-10'))
top_states_recent
```

Generate plot
```{r}
ggplot(top_states_recent, mapping =  aes(x=date, y = confirmed, 
                                         shape = top_states_recent$`Province/State`,
                                         color = top_states_recent$`Province/State`)) +
         geom_point() +
  scale_shape_manual(name = " ",
#labels = top_states_list$`Province/State`, 
                     values = c(4, 16, 17, 3, 7, 8, 11, 15)) +
  scale_color_manual(name = " ",
 #                    labels = top_states_list$`Province/State`, 
                     values = cbPalette) +
  labs(x="", y = "Confirmed", 
       title = "Confirmed COVID-19 Cases", 
       subtitle = "Top 8 US States", 
       caption = "Source: CDC and Johns Hopkins via GitHub")
```
